### 2022-01-08 cleave.tex
figures

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}.
Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of \acp{NCS}, namely the network itself.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{images/CLEAVE_NCS_structure.png}
    \caption{
        Structure of an emulated \ac{NCS} on top of CLEAVE.
        The State object implements the discrete-time evolution of the Plant.
        For this purpose, it holds two sets of special variables; Actuated Variables and Sensed Variables.
        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step.
        These are used to perform a time step update in the State, and update the values of the Sensed Variables.
        The values of the Sensed Variables are then, at the end of each time step, sampled and passed on to the Sensor objects, which in turn process them (for instance, to add noise) and send the processed values to the Controller.
        The Controller uses these values to calculate the new actuation commands and sends these to the Actuators.
        Finally, the Actuators process the actuation commands and hold the new values for the Actuated Variables until they are read at the next time step.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

>>>>

\todo[inline]{
    For the approach part, mention please also that it is open-source and where it can be found. Also, you keep the discussion around the implemented plant and the controller pretty short (and it is only showing up in the numerical part). Maybe it is better to discuss the details of the implemented plant and the controller already in Section 2 and mention that the intention is to build a liberary of plants and controllers in the future.
}

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.

@@

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
    \caption{
        Structure of an emulated \ac{NCS} on top of CLEAVE.\@
        The State object implements the discrete-time evolution of the Plant.
        For this purpose, it holds two sets of special variables; Actuated Variables and Sensed Variables.
        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step.
        These are used to perform a time step update in the State, and update the values of the Sensed Variables.
        The values of the Sensed Variables are then, at the end of each time step, sampled and passed on to the Sensor objects, which in turn process them (for instance, to add noise) and send the processed values to the Controller.
        The Controller uses these values to calculate the new actuation commands and sends these to the Actuators.
        Finally, the Actuators process the actuation commands and hold the new values for the Actuated Variables until they are read at the next time step.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

### 2022-01-08 conclusion.tex
figures

\section{Conclusion}\label{sec:conclusion}
\todo[inline]{TODO}

>>>>

\section{Conclusion}\label{sec:conclusion}

Performance characterization and evaluation of \acl{NCS} is hard, due to the tight interactions and cross-dependencies between network resource utilization and control-loop stability which complicate the scaling and repeatability of benchmarks.
This issue has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
To the best of our knowledge, \textcite{Zoppi2020NCSBench}'s work is the only available framework for \ac{NCS} benchmarking with repeatability and reproducibility in mind.

In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking.
Our framework, \ac{CLEAVE}, is fully parametrizable and built with industry-standard edge technologies and paradigms in mind.
We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well

% Benchmarking human-in-the-loop applications is hard, given
% their tight interaction with human users who complicate the scaling
% and repeatability of experiments. In this paper, we have presented
% a benchmarking suite for this type of applications, called Edge-
% Droid 1.0, capable of cutting out the need for users in performance
% evaluations. We achieve this by employing pre-recorded sensory
% input traces which we play over the network to the real applica-
% tion backend, employing a parameterized user model to react to
% feedback. We demonstrate its utility through a series of use case
% scenarios, from which we are able to extract metrics regarding la-
% tency both in regards to the application itself and the hardware
% stack. We believe the EdgeDroid 1.0 suite thus represents an impor-
% tant rst step towards enabling inexpensive and low-complexity
% large-scale research on the scaling limits of this type of applications,
% a requirement for wide adoption of the technology.

% Nonetheless, there is still future work to be done.
% The user model presented in this paper is only preliminary, and
% we are currently conducting research in characterizing user behav-
% ior when interacting with WCA applications in order to present
% a more complete model in the future. As mentioned in Section 4,
% in EdgeDroid 1.0, our model is that of a user who does not suer
% any of the shortcomings of real human users such as annoyance, fa-
% tigue, frustration, nausea. Rather, EdgeDroid 1.0 models a perfectly
% stoic user who is like an automaton and responds in a precisely
% reproducible and deterministic manner to the same system stimu-
% lus every time. Of course, no real human user is an automaton. In
% the future, we envision creating many versions of EdgeDroid (i.e.,
% EdgeDroid 2.0, EdgeDroid 3.0, etc.) that embody more human-like
% user models that more accurately emulate attributes such as those
% mentioned above. Experimental validation of these human-like user
% models via user studies will be an important part of our future work.
% We are also working on expanding the benchmarking suite to
% also work rst with other types of Wearable Cognitive Assistance,
% and later with other categories of human-in-the-loop applications.
% Other types of WCA we will consider in future iterations of the tool
% include real-time task-assistance WCA applications (such as the
% Ping-Pong application described in [7]), which don’t have a linear
% task model like task-guidance WCA and have tighter latency bounds
% and context- and information-providing WCA applications, for
% instance, applications which recognize faces and provide relevant
% social-media information related to that person. The latter also do
% not have a linear task model, but present more lax latency bounds.

### 2022-01-08 experiments.tex
figures

@@

The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk 2d physics library: \url{http://www.pymunk.org/en/latest/}}, executed at a constant \SI{120}{\hertz}.
For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.

@@

We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}.} were marked as ``toppled''.

Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \num{0}.}.

@@

\begin{figure}
    \centering
    \includegraphics[width=.8\columnwidth]{single_loop_toppled}
    \caption{Fraction of plants that toppled, per experimental setup.}%
    \label{fig:topple:fraction}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{single_loop_rtts}
        \caption{}\label{fig:single:rtt}
    \end{subfigure}\\%
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{single_loop_drop_frac}
        \caption{}\label{fig:single:drop}
    \end{subfigure}%
    \caption[caption]{Mean latency due to network and processing (\labelcref{fig:single:rtt}) for both single-loop scenarios, and packet losses (\labelcref{fig:single:drop}) for the single-loop scenario over the WiFi link.
    Error bars indicate \SI{95}{\percent} \acp{CI}.}%
    % \vspace{2pt}\raggedright{}\small\textsuperscript{a} The systematic offset of approx. \SI{2}{\milli\second} on scenarios with synthetic delay is due to an artifact of our implementation.
    \label{fig:single:network}
\end{figure}

@@

% As before, we start with a brief and simple analysis of plant stability.
\Cref{fig:video:stability} shows plots representing the fraction of repetitions with toppled penduli as well as the \ac{RMS} value of the angle of those penduli that did not topple.
\Cref{fig:video:network} shows network metrics for the scenario, with plots showing the amount of dropped \ac{UDP} datagrams and the average \acp{RTT}.

These results are interesting in their counter-intuitiveness.
Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:stability} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.

We can see an explanation for these behaviors in \cref{fig:video:network}.
Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).

@@

\begin{figure}
    \centering
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{video_topple_frac}
        \caption{}\label{fig:video:toppled}
    \end{subfigure}\\%
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{video_angle_rms}
        \caption{}\label{fig:video:rms}
    \end{subfigure}%
    \caption{
        Stability metrics for the realistic scenario.
        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
    }\label{fig:video:stability}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{video_drop_frac}
        \caption{}\label{fig:video:drop}
    \end{subfigure}\\%
    \begin{subfigure}[h]{\columnwidth}
        \centering
        \includegraphics[width=.8\columnwidth]{video_rtt}
        \caption{}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Network metrics for the realistic scenario.
        \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
        \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
    }\label{fig:video:network}
\end{figure}

>>>>

\todo[inline]{
    In the numerical part, you can gain a lot of space by reformatting the plots into environments with two or three plots next to each other. In addition, you need to comment more on Fig 4 (the key result for the single-loop baseline). The discussion of Fig 4 should take more space than the discussion of Fig 5.
}

@@

The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.

@@

We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \SI{0}{\percent}.}.

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_toppled}
        \caption{}%
        \label{fig:topple:fraction}
    \end{subfigure}%
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{single_loop_rms}
        \caption{}\label{fig:single:rms}
    \end{subfigure}%
    \caption[caption]{
        Stability metrics for the single-loop scenarios.
        \labelcref{fig:topple:fraction} shows the fraction of plants that toppled, per experimental setup.
        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
        }%
    \label{fig:single:network}
    \todo[inline]{More focus on this.}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{single_loop_rtts}
    \caption{
        Mean latency due to network and processing for both single-loop scenarios.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
    }\label{fig:single:rtt}
    \todo[inline]{less focus on this. mention packet drops are all below 0.2 percent}
\end{figure}

@@

\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
These results are interesting in their counter-intuitiveness.
Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.

\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.

We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[h]{.25\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{video_topple_frac}
        \caption{}\label{fig:video:toppled}
    \end{subfigure}%
    \begin{subfigure}[h]{.25\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{video_angle_rms}
        \caption{}\label{fig:video:rms}
    \end{subfigure}%
    \begin{subfigure}[h]{.25\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{video_drop_frac}
        \caption{}\label{fig:video:drop}
    \end{subfigure}%
    \begin{subfigure}[h]{.25\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{video_rtt}
        \caption{}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Results for the realistic scenario.
        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
        \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
        \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
        Error bars indicate \SI{95}{\percent} \ac{CI} in all plots.
    }\label{fig:video:results}
\end{figure*}

% \begin{figure}
%     \centering
%     \begin{subfigure}[h]{\columnwidth}
%         \centering
%         \includegraphics[width=.8\columnwidth]{video_drop_frac}
%         \caption{}\label{fig:video:drop}
%     \end{subfigure}\\%
%     \begin{subfigure}[h]{\columnwidth}
%         \centering
%         \includegraphics[width=.8\columnwidth]{video_rtt}
%         \caption{}\label{fig:video:rtt}
%     \end{subfigure}%
%     \caption{
%         Network metrics for the realistic scenario.
%         Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
%     }\label{fig:video:network}
% \end{figure}

### 2022-01-08 intro.tex
figures

\todo[inline, caption={}]{
    \begin{itemize}
        \item Still missing balance, and too long. Will fix towards the end.
    \end{itemize}
}

>>>>

\todo[inline, caption={}]{
    \begin{itemize}
        \item Still missing balance, and too long. Will fix towards the end.
        \item try to shorten the discussion around NCS. It is fine that you introduce it, but you don't need the extensive discussion. Likewise, for the experimental NCS works and the simulation works, you can package all that into a single paragraph, where your intent it simply to point out how people typically work in this domain. Then discuss the downsides, and lead over to mreo scalable and repeatable frameworks, i.e., NCSbench. Keep that discussion short and work out more strongly that it still heavily relies on the physical plant hardware (you just speak of hardware) being operated. The rest of the intro is fine.
    \end{itemize}

    CHANGE: removed 5Hz and 100ms cases from plots.
}
